Probability

### 2016-04-23 prob.tex
prob.tex \subsection{Quantum Probability Spaces} updated...

We now formalize a \emph{quantum probability space} as follows~\cite{BirkhoffVonNeumann1936,gleason1957,Redhead1987-REDINA,DBLP:journals/corr/abs-0910-2393,Maassen2010}.
We first assume an ambient Hilbert space~$\Hilb$ and define the
set of events $\qevents$ as all \emph{projections} on~$\Hilb$.
Each quantum state $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
induces a probability measure $\mu_{\psi}:\qevents\rightarrow[0,1]$
on the space of events defined for any event $E\in\qevents$ as follows\footnote{Recently, people extend the domain of $\mu_{\psi}$ to all operators~$\mathcal{A}$
on $\Hilb$ and consider $\mu_{\psi}:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.
When an operator $A\in\mathcal{A}$ is Hermitian, $\mu_{\psi}\left(A\right)$
is the expectation value of $A$. We does not take this approach because
we want to focus only on probability. }: 
\begin{equation}
\mu_{\psi}(E)= \ip{\psi}{E\psi} \label{eq:Born}
\end{equation}
We may verify our previous examples. When events are orthogonal and
commutative, they can be measured together and their probability is
additive: 
\[
\mu_{0}(\proj{0})+\mu_{0}(\proj{1})=\mu_{0}(\proj{+})+\mu_{0}(\proj{-})=1=\mu_{0}(\mathbb{1})\textrm{ .}
\]
On the other hand, $\proj{0}$ and $\proj{-}$ are non-commutative,
and their total probabilities are not a probability of an event
\begin{eqnarray*}
\mu_{0}(\proj{0})+\mu_{0}(\proj{-}) & = & \frac{3}{2}>1\textrm{ .}
\end{eqnarray*}
Similarly to the classical case, this probability measure must satisfy:
\begin{itemize}
\item $\mu(\mathbb{1})=1$, and 
\item for a collection of pairwise orthogonal $E_i$, we have
  $\mu(\sum_i E_i) = \sum_i \mu(E_i)$. 
\end{itemize}
\yutsung{There are three ways to charactize a quantum probability:
\begin{enumerate}
\item The above two conditions as Gleason did.
\item As a linear functional on a {*} algebra.
\item As induced by a state vector by the Born rule (\ref{eq:Born}).
\end{enumerate}
According to equation (37) in section 4 Quantum Probability in \cite{Maassen2010},
the equalivance between 1. and 2. only hold for $d\ge3$ because of
Gleason' theorem. Therefore, it seems hard to get insight for how
to handle Gleason's theorem by {*}-algebra formulism...

>>>>

We now formalize a \emph{quantum probability space} as follows~\cite{BirkhoffVonNeumann1936,gleason1957,Redhead1987-REDINA,DBLP:journals/corr/abs-0910-2393,Maassen2010}.
We first assume an ambient Hilbert space~$\Hilb$ and define the
set of events $\qevents$ as all \emph{projections} on~$\Hilb$.
Each quantum state $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
induces a probability measure $\mu_{\psi}:\qevents\rightarrow[0,1]$
on the space of events defined for any event $E\in\qevents$ as follows\footnote{Recently, people extend the domain of $\mu_{\psi}$ to all operators~$\mathcal{A}$
on $\Hilb$ and consider $\mu_{\psi}:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.
When an operator $A\in\mathcal{A}$ is Hermitian, $\mu_{\psi}\left(A\right)$
is the expectation value of $A$. We does not take this approach because
we want to focus only on probability. }: 
\begin{equation}
\mu_{\psi}(E)= \ip{\psi}{E\psi} \label{eq:Born}
\end{equation}
which is called the Born rule~\cite{Born1984,Mermin2007}.\\
\yutsung{One small thing is whether we should say
\begin{quote}
for each normalized state~$\ket{\psi}$...$\mu_{\psi}(E)=\ip{\psi}{E\psi}$
\end{quote}
as in Mermin~\cite{Mermin2007} and Maassen~\cite{Maassen2010}
or
\begin{quote}
for each state~$\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $...$\mu_{\psi}(E)=\frac{\ip{\psi}{E\psi}}{\ip{\psi}{\psi}}$
\end{quote}
as in Abramsky~\cite{DBLP:journals/corr/abs-0910-2393}. Personally,
I prefer Abramsky's approach, because normalizing a state in DQT is
messier than in CQT, and we have already considered states whose norm
is not one in our previous paper~\cite{DQT2014}.}\\
Our previous examples can be verified in terms of $\mu_{\psi}$. Because
$\proj{0}$ and $\proj{-}$ are non-commutative, their total probabilities
are not a probability of an event
\begin{eqnarray*}
\mu_{0}(\proj{0})+\mu_{0}(\proj{-}) & = & \frac{3}{2}>1\textrm{ .}
\end{eqnarray*}
When events are orthogonal, i.e., commutative, they can be measured
together and their probability is additive: 
\[
\mu_{0}(\proj{0})+\mu_{0}(\proj{1})=\mu_{0}(\proj{+})+\mu_{0}(\proj{-})=1=\mu_{0}(\mathbb{1})\textrm{ .}
\]
Similarly to the classical case, the total probability is one and
the orthogonal additivity can be summarized into the following two
rules:
\begin{itemize}
\item $\mu(\mathbb{1})=1$, and 
\item for a collection of pairwise orthogonal $E_{i}$, we have $\mu(\sum_{i}E_{i})=\sum_{i}\mu(E_{i})$.
\end{itemize}
These two conditions define quantum probability measures. Recall that
the Born rule (\ref{eq:Born}) induces a probability measure from
a state. Naturally, we may ask: is every probability measure induced
from a state by the Born rule? The answer is yes by Gleason's theorem
when the dimension~$\ge3$~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}.

\yutsung{There are three ways to charactize a quantum probability:
\begin{enumerate}
\item The above two conditions as Gleason did.
\item As a linear functional on a {*} algebra.
\item As induced by a state vector by the Born rule (\ref{eq:Born}).
\end{enumerate}
According to equation (37) in section 4 Quantum Probability in \cite{Maassen2010},
the equalivance between 1. and 2. only hold for $d\ge3$ because of
Gleason' theorem. Therefore, it seems hard to get insight for how
to handle Gleason's theorem by {*}-algebra formulism...
